Intrinsic Verification of Parsers and Formal Grammar Theory in Dependent Lambek Calculus (Extended Version)

### 2025-04-10 paper/paper.tex
copyright info filler
include publisher copyright info, journal info, and authors
ccs + keywords

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

@@

\newcommand{\simulsubst}[2]{#1\{#2\}}
\newcommand{\subst}[3]{\simulsubst {#1} {#2/#3}}
\newcommand{\el}{\mathsf{el}}
\newcommand{\letin}[3]{\mathsf{let}\, #1 = #2 \, \mathsf{in}\, #3}
\newcommand{\lamb}[2]{\lambda #1.\, #2}
\newcommand{\lamblto}[2]{\lambda^{{\lto}} #1.\, #2}
\newcommand{\lambtol}[2]{\lambda^{{\tol}} #1.\, #2}
\newcommand{\dlamb}[2]{{\lambda}^{{\&}} #1.\, #2}
\newcommand{\withlamb}[2]{{\lambda}^{{\&}} #1.\, #2}
\newcommand{\app}[2]{#1 \, #2}
\newcommand{\applto}[2]{#1 \, #2}
\newcommand{\apptol}[2]{#1 \mathop{{}^{\tol}} #2}
\newcommand{\PiTy}[3]{\textstyle\prod_{#1 : #2} #3}
\newcommand{\SigTy}[3]{\textstyle\sum_{#1 : #2} #3}
\newcommand{\PiTyLimit}[3]{\textstyle\prod\limits_{#1 : #2} #3}
\newcommand{\SigTyLimit}[3]{\textstyle\sum\limits_{#1 : #2} #3}
\newcommand{\LinPiTy}[3]{\textstyle\bigamp_{#1 : #2} #3}
\newcommand{\LinPiTyLimit}[3]{\bigwith\limits_{#1 : #2} #3}
\newcommand{\LinSigTy}[3]{\textstyle\bigoplus_{#1 : #2} #3}
\newcommand{\LinSigTyLimit}[3]{\bigoplus\limits_{#1 : #2} #3}
%% \newcommand{\DepWith}[2]{{\textstyle\bigamp}\limits_{#1}{#2}
%% \newcommand{\DepPlus}[2]{{\textstyle\bigoplus}\limits_{#1}{#2}
\newcommand{\GrTy}{\mathsf{Gr}}

@@

\author{Steven Schaefer}
\affiliation{\department{Electrical Engineering and Computer Science}
  \institution{University of Michigan}
  \country{USA}
}
\email{stschaef@umich.edu}
\orcid{0009-0007-1258-9501}

\author{Nathan Varner}
\affiliation{\department{Electrical Engineering and Computer Science}
  \institution{University of Michigan}
  \country{USA}
}
\email{nmvarner@umich.edu}
\orcid{0009-0000-3031-4930}

\author{Pedro H. Azevedo de Amorim}
\affiliation{
  \department{Department of Computer Science}
  \institution{University of Oxford}
  \country{UK}
}
\email{pedro.azevedo.de.amorim@cs.ox.ac.uk}
\orcid{0000-0002-8338-8973}

\author{Max S. New}
%% \authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{0000-0001-8141-195X}
\affiliation{
  \position{Assistant Professor}
  \department{Computer Science and Engineering}              %% \department is recommended
  \institution{University of Michigan}            %% \institution is required
  % \streetaddress{2260 Hayward St.}
  \city{Ann Arbor}
  \state{MI}
  % \postcode{48109-2121}
  \country{USA}                    %% \country is recommended
}
\email{maxsnew@umich.edu} %% \email is recommended

@@

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.

%% Keywords
%% comma separated list
\ifarxiv
\else
\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
\fi

>>>>

%%% The following is specific to PLDI '25 and the paper
%%% 'Intrinsic Verification of Parsers and Formal Grammar Theory in Dependent Lambek Calculus'
%%% by Steven Schaefer, Nathan Varner, Pedro Henrique Azevedo de Amorim, and Max S. New.
%%%
\setcopyright{cc}
\setcctype{by}
\acmDOI{10.1145/3729281}
\acmYear{2025}
\acmJournal{PACMPL}
\acmVolume{9}
\acmNumber{PLDI}
\acmArticle{178}
\acmMonth{6}
\received{2024-11-14}
\received[accepted]{2025-03-06}

@@

\newcommand{\simulsubst}[2]{#1\{#2\}}
\newcommand{\subst}[3]{\simulsubst {#1} {#2/#3}}
\newcommand{\el}{\mathsf{el}}
\newcommand{\letin}[3]{\mathsf{let}\, #1 = #2 \, \mathsf{in}\, #3}
\newcommand{\lamb}[2]{\lambda #1.\, #2}
\newcommand{\lamblto}[2]{\lambda^{{\lto}} #1.\, #2}
\newcommand{\lambtol}[2]{\lambda^{{\tol}} #1.\, #2}
\newcommand{\dlamb}[2]{{\lambda}^{{\&}} #1.\, #2}
\newcommand{\withlamb}[2]{{\lambda}^{{\&}} #1.\, #2}
\newcommand{\app}[2]{#1 \, #2}
\newcommand{\applto}[2]{#1 \, #2}
\newcommand{\apptol}[2]{#1 \mathop{{}^{\tol}} #2}
\newcommand{\PiTy}[3]{\textstyle\prod_{#1 : #2} #3}
\newcommand{\SigTy}[3]{\textstyle\sum_{#1 : #2} #3}
\newcommand{\PiTyLimit}[3]{\textstyle\prod\limits_{#1 : #2} #3}
\newcommand{\SigTyLimit}[3]{\textstyle\sum\limits_{#1 : #2} #3}
\newcommand{\LinPiTy}[3]{\textstyle\bigamp_{#1 : #2} #3}
\newcommand{\LinPiTyLimit}[3]{\bigwith\limits_{#1 : #2} #3}
\newcommand{\LinSigTy}[3]{\textstyle\bigoplus_{#1 : #2} #3}
\newcommand{\LinSigTyLimit}[3]{\bigoplus\limits_{#1 : #2} #3}
\newcommand{\GrTy}{\mathsf{Gr}}

@@

\author{Steven Schaefer}
\orcid{0009-0007-1258-9501}
\affiliation{%
  \institution{University of Michigan}
  \city{Ann Arbor}
  \country{USA}
}
\email{stschaef@umich.edu}

\author{Nathan Varner}
\orcid{0009-0000-3031-4930}
\affiliation{%
  \institution{University of Michigan}
  \city{Ann Arbor}
  \country{USA}
}
\email{nmvarner@umich.edu}

\author{Pedro Henrique Azevedo de Amorim}
\orcid{0000-0002-8338-8973}
\affiliation{%
  \institution{University of Oxford}
  \city{Oxford}
  \country{United Kingdom}
}
\email{pedro.azevedodeamorim@cs.ox.ac.uk}

\author{Max S. New}
\orcid{0000-0001-8141-195X}
\affiliation{%
  \institution{University of Michigan}
  \city{Ann Arbor}
  \country{USA}
}
\email{maxsnew@umich.edu}

@@

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003752.10010124.10010131.10010133</concept_id>
       <concept_desc>Theory of computation~Denotational semantics</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011050.10011017</concept_id>
       <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003766.10003771</concept_id>
       <concept_desc>Theory of computation~Grammars and context-free languages</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011041.10011688</concept_id>
       <concept_desc>Software and its engineering~Parsers</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011074.10011099.10011692</concept_id>
       <concept_desc>Software and its engineering~Formal software verification</concept_desc>
       <concept_significance>100</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[300]{Theory of computation~Denotational semantics}
\ccsdesc[300]{Software and its engineering~Domain specific languages}
\ccsdesc[500]{Theory of computation~Grammars and context-free languages}
\ccsdesc[500]{Software and its engineering~Parsers}
\ccsdesc[100]{Software and its engineering~Formal software verification}

%% Keywords
%% comma separated list
\ifarxiv
\else
\keywords{Linear types, Lambek calculus, Intrinsic verification}  %% \keywords are mandatory in final camera-ready submission
\fi
